---
title: "Social Media Data"
author: "Ryan Wesslen"
date: "19 Oct 2017"
output:
  iosp::ioslides_plus:
    box_colours:
      orange: "darkorange"
      slate: "slateblue2"
      yellow: ["yellow", "gold2"]
      grgold: ["springgreen4", header_text: "#FFD700"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rtweet); library(tidyverse)
```

## Overview

### Social Media Data {.col-10 .box .bg-green}

- How to Collect
- Querying Social Media
- Biases & Ethics

### Hands on: Twitter API {.col-10 .box .bg-yellow}

- RESTful: Current Tweets, Profile, Friends/Followers
- Streaming: Future Tweets
- "Big Data" Streaming

### Hands on: Facebook API {.col-10 .box .bg-orange}

- Searching public pages
- Scraping public pages
- Public page comments & likes

## Collecting Social Media Data {.build}

### Two Different Methods {.col-12 .box .bg-green}

- Web scraping: extract data from tagged source code of website
- Web APIs (application programming interface): use structured https requests that return JSON or XML files

### Two Types of APIs {.col-12 .box .bg-green}

- RESTful API's: queries for static data at current moment
- Streaming API's: changes in users' data in real time (future)

### Limitations {.col-12 .box .bg-green}

- API's limited to public data, subject to rate and historical limits
- Identify the best platform, **not** necessarily the easiest.

%end%

## Potential Issues {.vs1}

While there are opportunities in social media data, there are difficulties:

### Difficult Learning Curve {.bg-yellow .box-11}

>- API's & [data limitations](https://developer.twitter.com/en/docs/tweets/search/overview) 
>- [JSON](https://github.com/jimmoffitt/json2csv/blob/master/templates/tweet_standard.json) & [data formatting](https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/intro-to-tweet-json)
>- [Big data](http://www.bitbybitbook.com/en/observing-behavior/data/) & [scaling up](https://blog.insightdatascience.com/getting-started-with-aws-serverless-architecture-tutorial-on-kinesis-and-dynamodb-using-twitter-38a1352ca16d)
>- Self selection bias & [Representativeness](https://arxiv.org/pdf/1403.7400.pdf) 
>- Missing data (e.g., [geospatial](https://developer.twitter.com/en/docs/tutorials/tweet-geo-metadata))
>- [Querying data](https://developer.twitter.com/en/docs/tweets/data-dictionary/guides/tweet-timeline#filtering_tips) is harder than you think ([King, Lam, & Roberts, 2017](https://gking.harvard.edu/publications/computer-assisted-keyword-and-document-set-discovery-fromunstructured-text))
>- [Power laws](https://arxiv.org/abs/0706.1062) & 80/20 rule
>- [Bot or not bot](https://botometer.iuni.iu.edu/#!/)? The rise of [human-directed accounts](https://blog.twitter.com/official/en_us/topics/company/2017/Update-Russian-Interference-in-2016--Election-Bots-and-Misinformation.html#Human-Directed%20Accounts)
>- Unstructured data: text, images, urls, social network [GitHub](https://github.com/wesslen/summer2017-socialmedia)
>- Observational; difficult to run [experiments](http://www.bitbybitbook.com/en/running-experiments/) [(unless you're Zuck)](http://www.radiolab.org/story/trust-engineers/) 
>- [Ethics](http://www.bitbybitbook.com/en/ethics/ethics-intro/); role of IRB? [informed consent](http://www.wbur.org/cognoscenti/2014/07/15/experiment-facebook-steven-brykman)? [deletions](http://faculty.washington.edu/kstarbi/maddock_starbird_tweet_deletions.pdf)?

%end%

## Bots, ethics, and experimentation | Discussion

```{r, echo = FALSE}
#devtools::install_github("ijlyttle/vembedr")
vembedr::embed_youtube("u4NnZkxW3bY")
```

[Washington Post on Munger, 2017](https://www.washingtonpost.com/news/monkey-cage/wp/2016/11/17/this-researcher-programmed-bots-to-fight-racism-on-twitter-it-worked/)

## Note about third party research | Twitter Blog {.vcenter}

> **Studies of the impact of bots and automation on Twitter necessarily and systematically under-represent our enforcement actions because these defensive actions are not visible via our API**, and because they take place shortly after content is created and delivered via our streaming API. Furthermore, researchers using an API often overlook the substantial in-product features that prioritize the most relevant content. Based on user interests and choices, we limit the visibility of low-quality content using tools such as Quality Filter and Safe Search -- both of which are on by default for all of Twitter’s users and active for more than 97% of users.

Source: [Twitter Blog, Sept 28, 2017](https://blog.twitter.com/official/en_us/topics/company/2017/Update-Russian-Interference-in-2016--Election-Bots-and-Misinformation.html)

> “Research conducted by third parties through our search API about the impact of bots and misinformation on Twitter is almost always inaccurate and methodologically flawed,” Twitter said.

Source: [Washington Post, Sept 28, 2017](https://www.washingtonpost.com/business/technology/2017/09/27/32855bba-a3a0-11e7-ade1-76d061d56efa_story.html?utm_term=.51bd63e08c06)

## [`rtweet`](http://rtweet.info/articles/auth.html): REST and Streaming Twitter with R {.flexbox .vcenter}

- Open `01-rtweet-restapi.R` from `/01-rtweet/` folder.

- Based on the [rtweet authorization tutorial](http://rtweet.info/articles/auth.html).

## Get/Call Package | lines 4-7 {.vcenter}

```{r eval=FALSE}
# install from CRAN (uncomment out)
# install.packages("rtweet")

# load rtweet & tidyverse
library(rtweet); library(tidyverse)
```

## Create a Twitter API App {.build}

### Create Twitter app {.col-11 .box .bg-green}

- Go to [apps.twitter.com](https://apps.twitter.com/) and create a new app by providing a Name (rtweet_token), Description, and Website of your choosing.

### Callback URL {.col-11 .box .bg-green}

-  **Important**: use http://127.0.0.1:1410 for the Callback URL field.

### Accept Terms {.col-11 .box .bg-green}

- Check **yes** if you agree and then click “Create your Twitter application”.

## Run Examples {.vcenter}

- /01-rtweet/01-rtweet-restapi.R

- /01-rtweet/02-rtweet-streaming.R

- /01-rtweet/03-rtweet-streaming-large.R

- /02-Rfacebook/facebook-api.Rmd
