---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr); library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


## Tidyverse

### Example

Let's have a quick review of material.

```{r quiz}
quiz(
  question("What is tidyverse?",
    answer("R packages for machine learning only."),
    answer("R packages that automate (tidy) cleaning datasets."),
    answer("Collection of R packages under the tidy philosophy.", correct = TRUE),
    answer("A waste of time. Base R is the only true way to learn R.")
  ),
  question("What are the three tidy principles?", type = "multiple",
    answer("Each record is an observation.", correct = TRUE),
    answer("Each variable is its own column.", correct = TRUE),
    answer("Each value is a cell.", correct = TRUE),
    answer("Data should be stored in a way to minimize its size.",
           message = "See tidyr package website: http://tidyr.tidyverse.org/")
  ),
    question("What are core packages in tidyverse?", type = "multiple",
    answer("dplyr", correct = TRUE),
    answer("readr", correct = TRUE),
    answer("ggplot2", correct = TRUE),
    answer("purrr", correct = TRUE),
    answer("tidyr", correct = TRUE)
  ),
    question("What is piping (%>%)?", 
    answer("A function that overcomplicates data analysis."),
    answer("A simple way to link multiple functions, read left to right.", correct = TRUE),
    answer("A way to find out what inputs a function uses.")
  ),
   question("If you get an error message in R, what's the first thing you should do?",
    answer("Call 911"),
    answer("Sound the warning to Ryan (email, call, whatever it takes...)."),
    answer("Take a breath, copy the error, then Google it!", correct = TRUE),
    answer("Give up on R/programming and do everything in Excel.")
   )
)
```

### Run `tidyverse`

```{r call-tidyverse, exercise=TRUE}

```

This assumes you have the package `tidyverse`. Recall, what do you need to do if you do not have it?

## Data Wrangling Review

### Exercise with Hint

Use the mtcars (it's built in, no need to load) to run the following examples.

### Head function

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

How would you modify this to look at the first 8 rows?

```{r print-limit-2, exercise=TRUE}

```

```{r print-limit-hint-2}
head(mtcars, n = 8)
```

### Filter function

Use the filter() function to subset only includes records for cars with 4 or 6 cylinders.

```{r filter, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r filter-hint}
filter(mtcars, cyl %in% c("4","6"))
```

### Piping

Now rewrite that to use piping (%>%) and keep only the first five rows.

```{r filter2, exercise=TRUE}

```

```{r filter2-hint}
mtcars %>% 
  filter(cyl %in% c("4","6")) %>%
  head(n = 5)
```

### Group By

Next, using the piping operation, count the number of records by cylinder groups.

```{r groupby, exercise=TRUE}

```

```{r groupby-hint}
mtcars %>% 
  filter(cyl %in% c("4","6")) %>%
  group_by(cyl) %>%
  summarise(Count=n())
```

### Count

An even simpler way to rewrite the previous code is to use the count() function.

For example:

```{r count, exercise=TRUE}
filter(mtcars, cyl %in% c("4","6")) %>%
  count(cyl)
```

### dplyr functions

Let's review the dplyr functions.

```{r quiz2}
quiz(
  question("Which are the six dplyr functions?",
    answer("clean, import, subset, reduce, map, export"),
    answer("summarise, group_by, mutate, select, filter, arrange", correct = TRUE),
    answer("lm, rm, ls, post, pre, download"),
    answer("read_csv, write_csv, read_tsv, write_tsv, read_xls, write_xls")
  )
)